---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sparseVectors

The **sparseVectors** package implements an S4 class called `sparse_numeric'. This stores numeric vectors through keeping only the non zero values and positions, making it more efficient. 

## Features

The package supports:

### Core class
- `sparse_numeric` — an S4 class with slots:  
  - `value`: non-zero numeric values  
  - `pos`: positions of those values  
  - `length`: total length of the vector

### Arithmetic + math operations
- `sparse_add()` — elementwise addition  
- `sparse_sub()` — elementwise subtraction  
- `sparse_mult()` — elementwise multiplication  
- `sparse_crossprod()` — dot product  
- S4 overloads for `+`, `-`, `*`

### Statistics & transformations
- `mean()` method for sparse vectors (includes zeros)  
- `norm()` — Euclidean (L2) norm  
- `standardize()` — subtract mean and divide by standard deviation

### Coercion
- `as(numeric, "sparse_numeric")`  
- `as(sparse_numeric, "numeric")`

### Display
- Custom `show()` method  
- Custom `plot()` method for comparing two sparse vectors

## Installation

You can install the development version of sparseVectors from GitHub with:

``` r
# install.packages("pak")
pak::pak("Sreehana/sparseVectors")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(sparseVectors)
# Create a sparse_numeric vector
x <- as(c(1, 0, 2, 0), "sparse_numeric")
x
#> Sparse numeric vector of length 4
#> Non-zero values:
#>  pos value
#>    1     1
#>    3     2

# Compute the mean (including zeros)
mean(x)
#> 0.75

# Compute the Euclidean norm
norm(x)
#> sqrt(1^2 + 2^2) = 2.236

# Standardize the vector
standardize(x)

# Convert back to dense
as(x, "numeric")
#> 1 0 2 0
```
